---
title: "Kloosterman and Erbani et al., Cell 2024: Figure 1B"
author: "Daan J. Kloosterman"
date: "15/03/24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Figure 1B - Identifying cell types in the murine glioblastoma whole tumor scRNA-seq data (GEMM with PDG-Ink4a or PDG-p53 background)

```{r}
remotes::install_version("SeuratObject", "4.1.4", repos = c("https://satijalab.r-universe.dev", getOption("repos")))
remotes::install_version("Seurat", "4.4.0", repos = c("https://satijalab.r-universe.dev", getOption("repos")))
install.packages("Matrix", type = "source")
install.packages("irlba", type = "source")
library(Seurat)
library(SeuratObject)
library(Matrix)
library(irlba)
```

```{r}
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq"
setwd(work_dir)
# Set working directory to folder "'...'/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq" 

load("gbm-complete-subset.rds") 
DefaultAssay(gbm.combined_subset) <- "RNA"
```

# Manually annotated cell types based on cell specific markers

```{r}
## Assay is set to "RNA" as we will compare RNA expression between clusters
DefaultAssay(gbm.combined_subset) <- "RNA"
p3 <- DotPlot(gbm.combined_subset, group.by = "seurat_clusters", features = c("P2ry12", "Cd74", "Olig1" , "Gfap", "Pecam1", "Cd3e", "Itga4", "Mki67", "Des", "Ccr7", "Clec9a"))

p1 + p3
```


# Figure 1B: Annotation of unsupervised clusters as tumor cells, astrocytes, endothelial, T-cells, MDMs and MG based on the expression of cell specific markers, respectively Olig2, Gfap, Pecam1, Cd3e, Itga4 and P2ry12. Only PDG-Ink4a model is shown.

```{r}
## UMAP plot for Figure 1B
DimPlot(gbm.combined_subset, reduction = "umap", label = F)
```


# Split dataset for upcoming analyses and figures

```{r}
## Dataset is split into a macrophage and tumor cell dataset
gbm.myeloid <- gbm.combined_subset[, gbm.combined_subset$celltype %in% c("MG", "MDMs")]

DefaultAssay(gbm.myeloid) <- "integrated"
gbm.myeloid <- ScaleData(gbm.myeloid, verbose = FALSE)
gbm.myeloid <- RunPCA(gbm.myeloid, npcs = 30, verbose = FALSE)

gbm.myeloid <- RunUMAP(gbm.myeloid, reduction = "pca", dims = 1:25)
gbm.myeloid <- FindNeighbors(gbm.myeloid, reduction = "pca", dims = 1:25)
gbm.myeloid <- FindClusters(gbm.myeloid, resolution = 0.26, algorithm = 1)

## Prepping tumor cells
gbm.tumor  <- gbm.combined_subset[, gbm.combined_subset$celltype %in% c("Tumor cells")]

DefaultAssay(gbm.tumor) <- "integrated"
gbm.tumor <- ScaleData(gbm.tumor, verbose = FALSE)
gbm.tumor <- RunPCA(gbm.tumor, npcs = 30, verbose = FALSE)

gbm.tumor <- RunUMAP(gbm.tumor, reduction = "pca", dims = 1:30)
gbm.tumor <- FindNeighbors(gbm.tumor, reduction = "pca", dims = 1:30)
gbm.tumor <- FindClusters(gbm.tumor, resolution = 0.3, algorithm = 1)
```


# Saving data for downstream analyses

```{r}
save(gbm.myeloid,file="~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq/gbm-complete-myeloid.Rda")
save(gbm.tumor,file="~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq/gbm-complete-tumor.Rda")
```
