---
title: "Kloosterman and Erbani et al., Cell 2024: Figure 6A-B"
author: "Daan J. Kloosterman"
date: "21/03/24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Figure 6A-B: Analysing glioblastoma subtype lipid metabolic activity

# Loading packages/libraries required for the code
```{r}
library(Seurat)
library(nichenetr)
library(ggplot2)
library(scales)
```


# Setting working directory and loading all the pre-processed data files (only works when you press "Run Current Chunk")

```{r}
# Set working directory to folder "scrna-seq" 
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq"
setwd(work_dir)

load("gbm-complete-tumor_ink4a.Rda") # Model mostly used in manuscript
```


# Metabolic activity of cellular subtypes

```{r}
load("~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - code availibility/signatures/keggpathway_mmu.RData")

## Selecting only metabolic pathways
kg.mouse.met <- kg.mouse$kg.sets[kg.mouse$sigmet.idx] # Select which pathways

my_levels <- c("OPC","MES", "NPC", "AC")

## Prepping the data
fdscomplete <- c()
filtereddata <- gbm.tumor_ink4a  # Put here the dataset with clusterf you desire

# The script
for (i in 1:239){
  DefaultAssay(gbm.tumor_ink4a) <- "RNA"
  Gene <- unlist(as.vector(kg.mouse.met[[i]]))

##select the genes from the geneset that are in the Cluster.1 expression data
Gene <- Gene[Gene %in% rownames(gbm.tumor_ink4a)]

##Add score to every cell and create new seurat object. Not necessary to create a new object but in case anything goes wrong your original data is still there.
##This part is taken from the cellcycle classification, however I only used the part where Geneset scores were assigned to the cell
##create a score for every group
fdscores <- AddModuleScore(gbm.tumor_ink4a, features=list(c(Gene)), name="Gene",nbin=100)

# Average of fdscores
library(data.table)

gbm.tumor1 <- fdscores[, fdscores$Cell.Subtype.Grouped %in% c("OPC")] 
gbm.tumor2 <- fdscores[, fdscores$Cell.Subtype.Grouped %in% c( "MES")] 
gbm.tumor3 <- fdscores[, fdscores$Cell.Subtype.Grouped %in% c( "NPC")] 
gbm.tumor4 <- fdscores[, fdscores$Cell.Subtype.Grouped %in% c( "AC")] 


df1 <- mean(gbm.tumor1@meta.data$Gene) 
df2 <- mean(gbm.tumor2@meta.data$Gene) 
df3 <- mean(gbm.tumor3@meta.data$Gene) 
df4 <- mean(gbm.tumor4@meta.data$Gene) 


fdscompared <- as.data.frame(cbind(df1, df2, df3, df4))
fdscompared <- t(fdscompared)
row.names(fdscompared) <- my_levels

fdscomplete <- cbind(fdscomplete, fdscompared)
rm(fdscores)
}

  ## Finalizing data.frame and writing CSV file on desktop
colnames(fdscomplete) <- names(kg.mouse.met)

write.csv2(fdscomplete, "~/Desktop/GOscores_gbm-tumor_subtype_MOUSE.csv")
```


# Preparing data for radar plot

```{r}
## Load metabolic activity dataset
library(readxl)
GOscores_gbm_tumor_subtype_MOUSE <- as.data.frame(fdscomplete)


df <- cbind(GOscores_gbm_tumor_subtype_MOUSE$`mmu00061 Fatty acid biosynthesis`, GOscores_gbm_tumor_subtype_MOUSE$`mmu01212 Fatty acid metabolism`, GOscores_gbm_tumor_subtype_MOUSE$`mmu00100 Steroid biosynthesis`, GOscores_gbm_tumor_subtype_MOUSE$`mmu01040 Biosynthesis of unsaturated fatty acids`, GOscores_gbm_tumor_subtype_MOUSE$`mmu00020 Citrate cycle (TCA cycle)`, GOscores_gbm_tumor_subtype_MOUSE$`mmu00500 Starch and sucrose metabolism`,
            GOscores_gbm_tumor_subtype_MOUSE$`mmu00010 Glycolysis / Gluconeogenesis`,
            GOscores_gbm_tumor_subtype_MOUSE$`mmu00190 Oxidative phosphorylation`)
df <- as.data.frame(df)
row.names(df) <- GOscores_gbm_tumor_subtype_MOUSE$Samples
colnames(df) <- c("Fatty acid biosynthesis", "Fatty acid metabolism", "Steroid biosynthesis", "Biosynthesis of unsaturated fatty acids", "Citrate cycle (TCA cycle)", "Starch and sucrose metabolism", "Glycolysis", "Oxidative phosphorylation")

library(fmsb)

# Define the variable ranges: maximum and minimum
max_min <- data.frame(
  "Fatty.acid.biosynthesis" = c(0.76, -0.01), "Fatty.acid.metabolism" = c(0.76, -0.01), "Steroid.biosynthesis" = c(0.76, -0.01),
"Biosynthesis.of.unsaturated.fatty.acids" = c(0.76, -0.01),"Citrate.cycle.(TCA.cycle)" = c(0.76, -0.01), "Starch.and.sucrose.metabolism" = c(0.76, -0.01), "Glycolysis" = c(0.76, -0.01), "Oxidative phosphorylation" = c(0.76, -0.01))

rownames(max_min) <- c("Max", "Min")
colnames(max_min) <- c("Fatty acid biosynthesis", "Fatty acid metabolism", "Steroid biosynthesis", "Biosynthesis of unsaturated fatty acids","Citrate cycle (TCA cycle)",  "Starch and sucrose metabolism", "Glycolysis", "Oxidative phosphorylation")
```


# Figure 6A - Radar plot of relative metabolic activity of tumor cellular subtypes

```{r}
# Rescale value to range between 0-1
df_scaled <- round(apply(df, 2, scales::rescale), 2)
df_scaled <- as.data.frame(df_scaled)
head(df_scaled)

# Prepare the data for creating the radar plot using the fmsb package
# Variables summary
# Get the minimum and the max of every column  
col_max <- apply(df_scaled, 2, max)
col_min <- apply(df_scaled, 2, min)
# Calculate the average profile 
col_mean <- apply(df_scaled, 2, mean)
# Put together the summary of columns
col_summary <- t(data.frame(Max = col_max, Min = col_min, Average = col_mean))

# Produce radar plots showing both the average profile and the individual profile
# Bind variables summary to the data
df_scaled2 <- as.data.frame(rbind(col_summary, df_scaled))
head(df_scaled2)

opar <- par() 
# Define settings for plotting in a 3x4 grid, with appropriate margins:
par(mar = rep(0.8,4))
par(mfrow = c(1,4))
# Produce a radar-chart for each student
for (i in 4:nrow(df_scaled2)) {
  radarchart(
    df_scaled2[c(1:3, i), ],
    pfcol = c("#99999980",NA),
    pcol= c(NA,2), plty = 1, plwd = 2,
    title = row.names(df_scaled2)[i]
  )
}
# Restore the standard par() settings
par <- par(opar)
```

# Figure 6B - Dotplot of average expression of cholesterol import and biosynthesis genes per glioblastoma subtype

```{r}
# Figure 6B
DotPlot(gbm.tumor_ink4a, group.by = "Cell.Subtype", features = c("Vldlr" ,"Fdft1", "Dhcr24", "Hmgcr", "Cyp51", "Dhcr7"))
```
