---
title: "Figure 1I-J"
author: "Daan Kloosterman"
date: "30/07/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assessing the correlation between lipid-laden macrophages and the mesenchymal subtype enrichment in murine and human datasets

To further validate lipid-laden macrophages are enriched in mesenchymal glioblastoma dominant tumors, we will focus our analyses on these specific cellular subsets. Additionally, we probe an open source human glioblastoma scRNA-seq dataset to validate that this finding can be translated to patients.

# Loading packages/libraries required for the code

```{r}
library(Seurat)
library(ggplot2)
library(readxl)
library(nichenetr)
```


# Setting working directory and loading all the pre-processed data files (only works when you press "Run Current Chunk")

```{r}
# Set working directory to folder "pre-processed human scrna-seq" 
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq"
setwd(work_dir)

load("gbm-complete-tumor_polished.Rda")
load("gbm-complete-myeloid_polished.Rda")
```


# Figure 1I: Heatmap GPNMB-high signature

```{r}
Idents(gbm.myeloid) <- gbm.myeloid$clusterf
VlnPlot(gbm.myeloid, features = c("Gpnmb"))
new.cluster.ids <- c("GpnmbLow",	"GpnmbLow" , "GpnmbHigh",	"GpnmbLow",	"GpnmbLow",	"GpnmbLow", 	"GpnmbHigh",	 "GpnmbLow" )
names(new.cluster.ids) <- levels(gbm.myeloid)
gbm.myeloid <- RenameIdents(gbm.myeloid, new.cluster.ids)
gbm.myeloid$Gpnmb <- Idents(gbm.myeloid)


DimPlot(gbm.myeloid, reduction = "umap",  label = F, repel = T)
VlnPlot(gbm.myeloid, features = c("Gpnmb"), group.by = "clusterf", pt.size = 0)
# Finding classifiers for each subpopulation
Gpnmbmarkers <- FindAllMarkers(gbm.myeloid, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
Gpnmbmarkers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_log2FC)

# LLM assignment code ------------------------------------ ## 
Genes_high <- filter(Gpnmbmarkers, cluster == "GpnmbHigh")
Genes_low <- filter(Gpnmbmarkers, cluster == "GpnmbLow")

Gene1 <- Genes_high$gene[1:25] ## First LLM 
Gene2 <- Genes_low$gene[1:25] ## Now non LLM 
Gene_combined <- c(Gene2, Gene1)
gbm.myeloid <- ScaleData(gbm.myeloid)
Idents(gbm.myeloid) <- gbm.myeloid$Gpnmb
DoHeatmap(object = gbm.myeloid, features = Gene1)  + scale_fill_gradientn(colors = c("cornflowerblue", "white","brown1"))
```


# Figure 1J (1): Calculate % of GPNMB-high macrophages of total macrophages per sample, data generated from this analysis was plotted against the percentage of mesenchymal tumor cells (% of total tumor cells) which gathered from the data generated for Figure 1D.

```{r}
Idents(gbm.myeloid) <- gbm.myeloid$clusterf
DefaultAssay(gbm.myeloid) <- "RNA"
# Load GPNMBhigh signature as determined in supplementary figure 2a-d
Gene <- c("Gpnmb",  "Fabp5",  "Hmox1",  "Spp1" ,  "Arg1"  )

##Add score to every cell and create new seurat object. Not necessary to create a new object but in case anything goes wrong your original data is still there.
##This part is taken from the cellcycle classification, however I only used the part where Geneset scores were assigned to the cell
##create a score for every group
fdscores <- AddModuleScore(gbm.myeloid, features=list(c(Gene)), name="Gene",nbin=100)

## Calculate the average expression of the LLM signature by BMDMs and Microglia
gbm.myeloid_MDM3_GPNMB <- fdscores[, fdscores$clusterf %in% c("MDM3-GPNMB")]
gbm.myeloid_MG3_GPNMB <- fdscores[, fdscores$clusterf %in% c( "MG3-GPNMB")]
df1 <- median(gbm.myeloid_MDM3_GPNMB@meta.data$Gene1) 
df2 <- median(gbm.myeloid_MG3_GPNMB@meta.data$Gene1) 
averageLLM <- (df1 + df2)/2

##define the different groups in your Genelist, add '1' to every groupname. Groupnames can be checked in the metadata -> str(fdscores@meta.data)
groups <- c("Gene1")

##load function to create density values
densMode <- function(x){
  td <- density(x)
  tdx <- td$x
  tdy <- td$y
  minx <- tdx[which(diff(sign(diff(tdy)))==2)]
  peakx <- tdx[which(diff(sign(diff(tdy)))==-2)]
  return(list(minx=minx, maxy=peakx))
}

##create UMAPplots
DimPlot(gbm.myeloid, reduction = "umap", label = TRUE, pt.size = 0.5) 

##For every group determine the thesholds and plot several plots
for (i in groups){
  ##create densityplots and set cut-offs
  vl <- densMode(fdscores@meta.data[,i])[1]
  vl2 <- densMode(fdscores@meta.data[,i])[2]
  plot(density(fdscores@meta.data[,i]), main=paste("densityplot, GenesetScore, The higher the cell's score, the higher the Geneset's average expression ",i, sep=""))
  if(density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] > 6 || density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] <0.01){
    abline(v=vl$minx[2])
    threshold=vl$minx[2]
  }else{
    abline(v=vl$minx[1], col="red")
    threshold=vl$minx[1]
  }
  #plot(hist(fdscores@meta.data[,i], plot=F, freq=FALSE))
  #abline(h=500)
  #abline(v=vl2$maxy, col="red")
  
  ##classify the cells based on thresholds
  gbm.myeloid@meta.data[,paste("assignedto",i, sep="")] <- "nonclassified"
  gbm.myeloid@meta.data[which(fdscores@meta.data[,i]>averageLLM),paste("assignedto",i, sep="")] <- paste("assignedto",i, sep=" ") ## Use 0.4 for MG
  
  ##plot figures
  plot(DimPlot(gbm.myeloid, pt.size=1,group.by=paste("assignedto",i, sep=""))+labs(title = paste("assignment based on cutoff densityplot ",i, sep=" ")))
  
  plot(FeaturePlot(fdscores,pt.size=1, features=i))
  plot(VlnPlot(fdscores, features=i, group.by="clusterf", pt.size = 0.00001)) + ggtitle("Pseudopalisading Necrosis")
  plot(VlnPlot(fdscores, features=i, group.by= c("samplename"), pt.size = 0))
  DimPlot(fdscores, reduction = "umap", group.by = c("clusterf"), label = TRUE, pt.size = 0.5) 
}
Idents(gbm.myeloid) <- gbm.myeloid@meta.data$assignedtoGene1
new.cluster.ids <- c("non-LLMs", "LLMs")

names(new.cluster.ids) <- levels(gbm.myeloid)
gbm.myeloid <- RenameIdents(gbm.myeloid, new.cluster.ids)
gbm.myeloid$LLM <- Idents(gbm.myeloid)
DimPlot(gbm.myeloid, reduction = "umap",  group.by = "LLM", label = F)

## Print
x_abundance <- t(table(gbm.myeloid@meta.data$assignedtoGene1, gbm.myeloid$samplename))
write.csv(x_abundance, file = "~/Desktop/x_abundance_LipidLaden-mouse.csv")
```


# Loading patient data

```{r}
# Set working directory to folder "pre-processed human scrna-seq" 
work_dir <- "~/surfdrive/Code Availability/Kloosterman et al. 2024 - data availibility/GEO Data - processed/pre-processed human scrna-seq"
setwd(work_dir)

# HAM737
data_dir <- paste(work_dir, "/HAM737/", sep = "")
list.files(data_dir) # Should show barcodes.tsv, genes.tsv, and matrix.mtx
expression_matrix <- Read10X(data.dir = data_dir)
HAM737 = CreateSeuratObject(counts = expression_matrix)
HAM737 <- NormalizeData(HAM737, verbose = FALSE)
HAM737 <- FindVariableFeatures(HAM737, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
HAM737$signature <- c("HAM737")

# HAM738
data_dir <- paste(work_dir, "/HAM738/", sep = "")
list.files(data_dir) # Should show barcodes.tsv, genes.tsv, and matrix.mtx
expression_matrix <- Read10X(data.dir = data_dir)
HAM738 = CreateSeuratObject(counts = expression_matrix)
HAM738 <- NormalizeData(HAM738, verbose = FALSE)
HAM738 <- FindVariableFeatures(HAM738, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
HAM738$signature <- c("HAM738")

# opk356b
data_dir <- paste(work_dir, "/opk356b/outs/filtered_gene_bc_matrices/GRCh38/", sep = "")
list.files(data_dir) # Should show barcodes.tsv, genes.tsv, and matrix.mtx
expression_matrix <- Read10X(data.dir = data_dir)
opk356b = CreateSeuratObject(counts = expression_matrix)
opk356b <- NormalizeData(opk356b, verbose = FALSE)
opk356b <- FindVariableFeatures(opk356b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk356b$signature <- c("opk356b")

# opk363b
opk363b <- Read10X_h5("/opk363b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk363b = CreateSeuratObject(counts = opk363b)
opk363b <- NormalizeData(opk363b, verbose = FALSE)
opk363b <- FindVariableFeatures(opk363b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk363b$signature <- c("opk363b")

# opk363NC
data_dir <- paste(work_dir, "/opk363NC/", sep = "")
list.files(data_dir) # Should show barcodes.tsv, genes.tsv, and matrix.mtx
expression_matrix <- Read10X(data.dir = data_dir)
opk363NC = CreateSeuratObject(counts = expression_matrix)
opk363NC <- NormalizeData(opk363NC, verbose = FALSE)
opk363NC <- FindVariableFeatures(opk363NC, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk363NC$signature <- c("opk363NC")

# opk364b
opk364b <- Read10X_h5("/opk364b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk364b = CreateSeuratObject(counts = opk364b)
opk364b <- NormalizeData(opk364b, verbose = FALSE)
opk364b <- FindVariableFeatures(opk364b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk364b$signature <- c("opk364b")

# opk368b
opk368b <- Read10X_h5("/opk368b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk368b = CreateSeuratObject(counts = opk368b)
opk368b <- NormalizeData(opk368b, verbose = FALSE)
opk368b <- FindVariableFeatures(opk368b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk368b$signature <- c("opk368b")

# opk369b
opk369b <- Read10X_h5("/opk369b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk369b = CreateSeuratObject(counts = opk369b)
opk369b <- NormalizeData(opk369b, verbose = FALSE)
opk369b <- FindVariableFeatures(opk369b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk369b$signature <- c("opk369b")

# opk373b
opk373b <- Read10X_h5("/opk373b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk373b = CreateSeuratObject(counts = opk373b)
opk373b <- NormalizeData(opk373b, verbose = FALSE)
opk373b <- FindVariableFeatures(opk373b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk373b$signature <- c("opk373b")

# opk374b
opk374b <- Read10X_h5("/opk374b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk374b = CreateSeuratObject(counts = opk374b)
opk374b <- NormalizeData(opk374b, verbose = FALSE)
opk374b <- FindVariableFeatures(opk374b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk374b$signature <- c("opk374b")

# opk377b
opk377b <- Read10X_h5("/opk377b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk377b = CreateSeuratObject(counts = opk377b)
opk377b <- NormalizeData(opk377b, verbose = FALSE)
opk377b <- FindVariableFeatures(opk377b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk377b$signature <- c("opk377b")

# opk380b
opk380b <- Read10X_h5("/opk380b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk380b = CreateSeuratObject(counts = opk380b)
opk380b <- NormalizeData(opk380b, verbose = FALSE)
opk380b <- FindVariableFeatures(opk380b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk380b$signature <- c("opk380b")

# opk389A
opk389A <- Read10X_h5("/opk389A/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk389A = CreateSeuratObject(counts = opk389A)
opk389A <- NormalizeData(opk389A, verbose = FALSE)
opk389A <- FindVariableFeatures(opk389A, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk389A$signature <- c("opk389A")

# opk389b
opk389b <- Read10X_h5("/opk389b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk389b = CreateSeuratObject(counts = opk389b)
opk389b <- NormalizeData(opk389b, verbose = FALSE)
opk389b <- FindVariableFeatures(opk389b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk389b$signature <- c("opk389b")

# opk390b
opk390b <- Read10X_h5("/opk390b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk390b = CreateSeuratObject(counts = opk390b)
opk390b <- NormalizeData(opk390b, verbose = FALSE)
opk390b <- FindVariableFeatures(opk390b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk390b$signature <- c("opk390b")

# opk396b
opk396b <- Read10X_h5("/opk396b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk396b = CreateSeuratObject(counts = opk396b)
opk396b <- NormalizeData(opk396b, verbose = FALSE)
opk396b <- FindVariableFeatures(opk396b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk396b$signature <- c("opk396b")

# opk397b
opk397b <- Read10X_h5("/opk397b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk397b = CreateSeuratObject(counts = opk397b)
opk397b <- NormalizeData(opk397b, verbose = FALSE)
opk397b <- FindVariableFeatures(opk397b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk397b$signature <- c("opk397b")

# opk400b
opk400b <- Read10X_h5("/opk400b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk400b = CreateSeuratObject(counts = opk400b)
opk400b <- NormalizeData(opk400b, verbose = FALSE)
opk400b <- FindVariableFeatures(opk400b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk400b$signature <- c("opk400b")

# opk402b
opk402b <- Read10X_h5("/opk402b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk402b = CreateSeuratObject(counts = opk402b)
opk402b <- NormalizeData(opk402b, verbose = FALSE)
opk402b <- FindVariableFeatures(opk402b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk402b$signature <- c("opk402b")

# opk407b
opk407b <- Read10X_h5("/opk407b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk407b = CreateSeuratObject(counts = opk407b)
opk407b <- NormalizeData(opk407b, verbose = FALSE)
opk407b <- FindVariableFeatures(opk407b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk407b$signature <- c("opk407b")

# opk408b
opk408b <- Read10X_h5("/opk408b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk408b = CreateSeuratObject(counts = opk408b)
opk408b <- NormalizeData(opk408b, verbose = FALSE)
opk408b <- FindVariableFeatures(opk408b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk408b$signature <- c("opk408b")

# opk409b
opk409b <- Read10X_h5("/opk409b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk409b = CreateSeuratObject(counts = opk409b)
opk409b <- NormalizeData(opk409b, verbose = FALSE)
opk409b <- FindVariableFeatures(opk409b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk409b$signature <- c("opk409b")

# opk419b
opk419b <- Read10X_h5("/opk419b/filtered_feature_bc_matrix.h5", use.names = TRUE, unique.features = TRUE)
opk419b = CreateSeuratObject(counts = opk419b)
opk419b <- NormalizeData(opk419b, verbose = FALSE)
opk419b <- FindVariableFeatures(opk419b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk419b$signature <- c("opk419b")

# opk420b
opk420b <- Read10X_h5("/opk420b/filtered_gene_bc_matrices_h5.h5", use.names = TRUE, unique.features = TRUE)
opk420b = CreateSeuratObject(counts = opk420b)
opk420b <- NormalizeData(opk420b, verbose = FALSE)
opk420b <- FindVariableFeatures(opk420b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk420b$signature <- c("opk420b")

# opk426b
opk426b <- Read10X_h5("/opk426b/filtered_feature_bc_matrix.h5", use.names = TRUE, unique.features = TRUE)
opk426b = CreateSeuratObject(counts = opk426b)
opk426b <- NormalizeData(opk426b, verbose = FALSE)
opk426b <- FindVariableFeatures(opk426b, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
opk426b$signature <- c("opk426b")


```


# Processing data

```{r}
combined <- c(HAM738, opk356b, opk363NC, opk363b, opk364b, opk368b, opk369b,opk373b, opk374b, opk377b, opk380b, opk389A, opk389b, opk390b, opk396b, opk397b, opk400b, opk402b, opk407b, opk408b, opk409b, opk419b, opk420b)

seurat_object <- merge(opk356b, y = combined_2)
seurat_object <- subset(seurat_object, subset = nFeature_RNA > 200 & nFeature_RNA < 4000)
seurat_object <- NormalizeData(seurat_object, verbose = FALSE)
seurat_object <- FindVariableFeatures(seurat_object, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
seurat_object <- ScaleData(seurat_object, verbose = FALSE)
seurat_object <- RunPCA(seurat_object, npcs = 30, verbose = FALSE)
seurat_object <- FindNeighbors(seurat_object, dims = 1:10)
seurat_object <- FindClusters(seurat_object, algorithm = 1, random.seed = 200, resolution = 0.58)
seurat_object <- RunUMAP(seurat_object, dims = 1:10)
DimPlot(seurat_object, reduction = "umap",  group.by = "seurat_clusters", label = TRUE)
combinedHuman10x_merged <- seurat_object
```


# Saving data before starting the analysis (takes quite some time to create this file) to prevent the time-consuming preprocessing steps in the future

```{r}
setwd("~/Users/d.kloosterman/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq") # Set working directory to folder where integrated and filtered data will be saved: "scrna-seq"

save(combinedHuman10x_merged,file="combinedHUman10x_merged.Rda")
```


# Setting working directory and loading all the pre-processed data files (only works when you press "Run Current Chunk")

```{r}
# Set working directory to folder "pre-processed human scrna-seq" 
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq"
setwd(work_dir)

load("combinedHuman10x_merged.Rda")
```


# Figure 1J (2-1): Calculate % of GPNMB-high macrophages of total macrophages per patients, data generated from this analysis was plotted against the percentage of mesenchymal tumor cells (% of total tumor cells) which is generated in the next chunk.

```{r}
# Select macrophage clusters
filtereddata <- combinedHuman10x_merged[, combinedHuman10x_merged$seurat_clusters %in% c("0", "1", "13", "6", "15", "24", "17")]
filtereddata <- FindNeighbors(filtereddata, dims = 1:30)
filtereddata <- FindClusters(filtereddata, algorithm = 2, random.seed = 200, resolution = 0.1)
filtereddata <- RunUMAP(filtereddata, dims = 1:30)
DimPlot(filtereddata, reduction = "umap",  group.by = "seurat_clusters", label = TRUE)

library(RColorBrewer)

## Lipid-laden macrophage signature
Gene <- c("GPNMB",  "FABP5",  "HMOX1",  "SPP1" ,  "ARG1" )

FeaturePlot(filtereddata, features = Gene)

##select the genes from the geneset that are in the Cluster.1 expression data
Gene <- Gene[Gene %in% rownames(filtereddata)]

##Add score to every cell and create new seurat object. Not necessary to create a new object but in case anything goes wrong your original data is still there.
##This part is taken from the cellcycle classification, however I only used the part where Geneset scores were assigned to the cell
##create a score for every group
fdscores <- AddModuleScore(filtereddata, features=list(c(Gene)), name="Gene",nbin=100)

##define the different groups in your Genelist, add '1' to every groupname. Groupnames can be checked in the metadata -> str(fdscores@meta.data)
groups <- c("Gene1")

VlnPlot(fdscores, features=i, group.by="seurat_clusters", pt.size = 0)

gbm.human <- fdscores[, fdscores$seurat_clusters %in% c("1")] # Cluster with highest score
averageLLM <- median(gbm.human@meta.data$Gene1) 

##load function to create density values
densMode <- function(x){
  td <- density(x)
  tdx <- td$x
  tdy <- td$y
  minx <- tdx[which(diff(sign(diff(tdy)))==2)]
  peakx <- tdx[which(diff(sign(diff(tdy)))==-2)]
  return(list(minx=minx, maxy=peakx))
}

##create UMAPplots
DimPlot(filtereddata, reduction = "umap", label = TRUE, pt.size = 0.5) 

##For every group determine the thesholds and plot several plots
for (i in groups){
  ##create densityplots and set cut-offs
  vl <- densMode(fdscores@meta.data[,i])[1]
  vl2 <- densMode(fdscores@meta.data[,i])[2]
  plot(density(fdscores@meta.data[,i]), main=paste("densityplot, GenesetScore, The higher the cell's score, the higher the Geneset's average expression ",i, sep=""))
  if(density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] > 6 || density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] <0.01){
    abline(v=vl$minx[2])
    threshold=vl$minx[2]
  }else{
    abline(v=vl$minx[1], col="red")
    threshold=vl$minx[1]
  }
  #plot(hist(fdscores@meta.data[,i], plot=F, freq=FALSE))
  #abline(h=500)
  #abline(v=vl2$maxy, col="red")
  
  ##classify the cells based on thresholds
  filtereddata@meta.data[,paste("assignedto",i, sep="")] <- "nonclassified"
  filtereddata@meta.data[which(fdscores@meta.data[,i]>averageLLM),paste("assignedto",i, sep="")] <- paste("assignedto",i, sep=" ")
5
  ##plot figures
  plot(DimPlot(filtereddata, pt.size=1,group.by=paste("assignedto",i, sep=""))+labs(title = paste("assignment based on cutoff densityplot ",i, sep=" ")))
  
  plot(FeaturePlot(fdscores,pt.size=1, features=i))
  plot(VlnPlot(fdscores, features=i, group.by="signature", pt.size = ))
}

## Use this to calculate the relative abundance of lipid-laden macrophages (assigned to Gene1) per patient 
x_abundance <- t(table(filtereddata@meta.data$assignedtoGene1, filtereddata$signature))

mean_LLM <- aggregate(fdscores$Gene1, list(fdscores$signature), FUN=mean) 

filtereddata$LLM <- filtereddata@meta.data$assignedtoGene1

write.csv(x_abundance, file = "~/Desktop/x_abundance_LipidLaden-patient.csv")
write.csv(mean_LLM, file = "~/Desktop/x_mean_LipidLaden-signature-patient.csv")
```


# Figure 1J (2-1): Calculate glioblastoma cellular subtype distribution per patient, data generated from this analysis was plotted against the percentage of GPNMB-high macrophages (% of total macrophages) which is generated in the previous chunk.

```{r}
# Select tumor clusters
filtereddata <- combinedHuman10x_merged[, combinedHuman10x_merged$seurat_clusters %in% c("5", "20","7", "14", "20", "10", "4", "11", "19", "23", "12")]
filtereddata <- FindNeighbors(filtereddata, dims = 1:5)
filtereddata <- FindClusters(filtereddata, algorithm = 2, random.seed = 200, resolution = 0.1)
filtereddata <- RunUMAP(filtereddata, dims = 1:5)
DimPlot(filtereddata, reduction = "umap",  group.by = "seurat_clusters", label = TRUE)


# Loading glioblastoma cellular subtype modules (Neftel et al., 2019)
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - code availibility/signatures"
setwd(work_dir)
Cell_Signatures <- read_excel("Glioblastoma_meta-module_genelist.xlsx")

MES1.features <- Cell_Signatures$MES1
MES2.features <- Cell_Signatures$MES2
AC.features <- Cell_Signatures$AC
OPC.features <- Cell_Signatures$OPC
NPC1.features <- Cell_Signatures$NPC1
NPC2.features <- Cell_Signatures$NPC2

## Function to assign cellular subtype
assignSubtype <- function (object, MES1.features,  MES2.features, AC.features, OPC.features, NPC1.features, NPC2.features ,set.ident = FALSE, ctrl_genes = 100,
                           ...)
{
  name <- "Cell.Subtype"
  features <- list(  MES1.Score = MES1.features, MES2.Score =  MES2.features, AC.Score = AC.features, OPC.Score = OPC.features, NPC1.Score = NPC1.features, NPC2.Score = NPC2.features)
  object.cc <- AddModuleScore(object = object,
                              features = features,
                              name = name,
                              ctrl = min(c(vapply(X = features, FUN = length, FUN.VALUE = numeric(length = 1))), 
                                         ctrl_genes), 
                              ...)
  cc.columns <- grep(pattern = name, x = colnames(x = object.cc[[]]),
                     value = TRUE)
  cc.scores <- object.cc[[cc.columns]]
  rm(object.cc)
  CheckGC()
  assignments <- apply(X = cc.scores, MARGIN = 1, FUN = function(scores,
                                                                 first = "MES1", second = "MES2", third = "AC", fourth = "OPC", fifth = "NPC1", sixth = "NPC2", null = "Undecided") {
    if (all(scores < -0)) {
      return(null)
    }
    else {
      return(c(first, second, third, fourth, fifth, sixth)[which(x = scores == max(scores))])
    }
  }
  )
  cc.scores <- merge(x = cc.scores, y = data.frame(assignments),
                     by = 0)
  colnames(x = cc.scores) <- c("rownames", "MES1.Score", "MES2.Score", "AC.Score","OPC.Score", "NPC1.Score", "NPC2.Score",
                               "Cell.Subtype")
  rownames(x = cc.scores) <- cc.scores$rownames
  cc.scores <- cc.scores[, c("MES1.Score", "MES2.Score", "AC.Score","OPC.Score", "NPC1.Score", "NPC2.Score","Cell.Subtype")]
  object[[colnames(x = cc.scores)]] <- cc.scores
  if (set.ident) {
    object[["old.ident"]] <- Idents(object = object)
    Idents(object = object) <- "Cell.Subtype"
  }
  return(object)
}

filtereddata <- assignSubtype(filtereddata, MES1.features = MES1.features, MES2.features = MES2.features, AC.features = AC.features, OPC.features = OPC.features, NPC1.features = NPC1.features, NPC2.features = NPC2.features)

# Plot cellular celltype distribution
DimPlot(filtereddata, reduction = "umap", group.by = "Cell.Subtype", pt.size = 2)

# Create .csv file with subtype composition per samples
x_abundance <- t(table(filtereddata$Cell.Subtype, filtereddata$signature))
write.csv(x_abundance, file = "~/Desktop/x_abundance_Cell.Subtype-patient.csv")
```
