---
title: "Kloosterman and Erbani et al., Cell 2024: Figure 1C-D"
author: "Daan J. Kloosterman"
date: "15/03/24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Figure 1C-D - Analysing glioblastoma tumor cellular subtype heterogeneity in GEMM with PDG-Ink4a and PDG-shp53 background

This modules contains the code to generate Figure 1C and the data to make Figure 1D, which includes a function to assign an unbiased glioblastoma cellular subtype identity and its pseudolocation to each individual tumor cell. Here, we assign tumor cell classifications which is later used in downstream analyses in the manuscript.


# Loading packages/libraries required for the code

```{r}
library(Seurat)
library(nichenetr) 
library(ggplot2)
```


# Setting working directory and loading all the pre-processed data files (only works when you press "Run Current Chunk")

```{r}
# Set working directory to folder "scrna-seq" 
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq"
setwd(work_dir)
load("gbm-complete-tumor.Rda")
load("gbm-complete-subset.Rda")
```


# Cleaning up the selected tumor cells from the combined dataset - PDG-Ink4a model

```{r}
## Plotting the tumor cell dataset
DimPlot(gbm.tumor, reduction = "umap", group.by = "seurat_clusters")

# Split tumor dataset per genetic background for comparisons between primary and recurrent state
gbm.tumor_ink4a <- gbm.tumor[, gbm.tumor$samplename %in% c("ink4a_primary_CD45_b", "ink4a_primary_CD45_c", "ink4a_primary_CD45_d", "ink4a_recurrent_CD45_b", "ink4a_recurrent_CD45_c", "ink4a_recurrent_CD45_d")]

# Remove doublets with other cell types (AP2ry12 = MG, Aif1 = MDMs, Cd3e = T cells and Pecam1 = Endothelial cell specific)
DefaultAssay(gbm.tumor_ink4a) <- "RNA"
fdscores <- AddModuleScore(gbm.tumor_ink4a, features=list(c("P2ry12","Aif1", "Cd3e", "Pecam1", "Des", "Gfap", "Ccr7", "Clec4e")), name="Gene",nbin=100)

##define the different groups in your Genelist, add '1' to every groupname. Groupnames can be checked in the metadata -> str(fdscores@meta.data)
groups <- c("Gene1")

##load function to create density values
densMode <- function(x){
  td <- density(x)
  tdx <- td$x
  tdy <- td$y
  minx <- tdx[which(diff(sign(diff(tdy)))==2)]
  peakx <- tdx[which(diff(sign(diff(tdy)))==-2)]
  return(list(minx=minx, maxy=peakx))
}

## A density plot is made of the fds score, everything above a score of 0 is considered a doublet with another cell type
for (i in groups){
  ##create densityplots and set cut-offs
  vl <- densMode(fdscores@meta.data[,i])[1]
  vl2 <- densMode(fdscores@meta.data[,i])[2]
  plot(density(fdscores@meta.data[,i]), main=paste("densityplot, GenesetScore, The higher the cell's score, the higher the Geneset's average expression ",i, sep=""))
  if(density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] > 6 || density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] <0.01){
    abline(v=vl$minx[2])
    threshold=vl$minx[2]
  }else{
    abline(v=vl$minx[1], col="red")
    threshold=vl$minx[1]
  }
  #plot(hist(fdscores@meta.data[,i], plot=F, freq=FALSE))
  #abline(h=500)
  #abline(v=vl2$maxy, col="red")
  
  ##classify the cells based on thresholds of 0.1
  gbm.tumor_ink4a@meta.data[,paste("assignedto",i, sep="")] <- "nonclassified"
  gbm.tumor_ink4a@meta.data[which(fdscores@meta.data[,i]>0),paste("assignedto",i, sep="")] <- paste("assignedto",i, sep=" ")
}

## Doublets are removed
Idents(gbm.tumor_ink4a) <- gbm.tumor_ink4a@meta.data$assignedtoGene1
new.cluster.ids <- c( "Doublets", "Tumor cells")
names(new.cluster.ids) <- levels(gbm.tumor_ink4a)
gbm.tumor_ink4a <- RenameIdents(gbm.tumor_ink4a, new.cluster.ids)
gbm.tumor_ink4a$celltype <- Idents(gbm.tumor_ink4a)
DimPlot(gbm.tumor_ink4a, reduction = "umap", group.by = "celltype")

## Clean up the data by removing doublets cells 
gbm.tumor_ink4a <- gbm.tumor_ink4a[, gbm.tumor_ink4a$celltype %in% c("Tumor cells")]

DefaultAssay(gbm.tumor_ink4a) <- "integrated"
gbm.tumor_ink4a <- RunUMAP(gbm.tumor_ink4a, reduction = "pca", dims = 1:30)
gbm.tumor_ink4a <- FindNeighbors(gbm.tumor_ink4a, reduction = "pca", dims = 1:30)
gbm.tumor_ink4a <- FindClusters(gbm.tumor_ink4a, resolution = 0.3, algorithm = 1)

## Plotting the tumor cell dataset and removing large outlier cluster (4)
DimPlot(gbm.tumor_ink4a, reduction = "umap", group.by = "seurat_clusters")
gbm.tumor_ink4a <- gbm.tumor_ink4a[, gbm.tumor_ink4a$seurat_clusters %in% c("0", "1", "2", "3", "5")]
DefaultAssay(gbm.tumor_ink4a) <- "integrated"
gbm.tumor_ink4a <- RunUMAP(gbm.tumor_ink4a, reduction = "pca", dims = 1:30)

DimPlot(gbm.tumor_ink4a, reduction = "umap", group.by = "stage_model")
```


# Cleaning up the selected tumor cells from the combined dataset - PDG-shp53 model

```{r}
## Plotting the tumor cell dataset
# gbm.tumor <- SCTransform(gbm.tumor)
DimPlot(gbm.tumor, reduction = "umap", group.by = "seurat_clusters")

# Split tumor dataset per genetic background for comparisons between primary and recurrent state
gbm.tumor_p53 <- gbm.tumor[, gbm.tumor$samplename %in% c( "p53_primary_CD45_a", "p53_primary_CD45_b", "p53_primary_CD45_c", "p53_primary_CD45_d", "p53_recurrent_CD45_a", "p53_recurrent_CD45_b", "p53_recurrent_CD45_c")]


# Remove doublets with other cell types (AP2ry12 = MG, Aif1 = MDMs, Cd3e = T cells and Pecam1 = Endothelial cell specific)
DefaultAssay(gbm.tumor_p53) <- "RNA"
fdscores <- AddModuleScore(gbm.tumor_p53, features=list(c("P2ry12","Aif1", "Cd3e", "Pecam1", "Des", "Gfap", "Ccr7", "Clec4e")), name="Gene",nbin=100)

                           
##define the different groups in your Genelist, add '1' to every groupname. Groupnames can be checked in the metadata -> str(fdscores@meta.data)
groups <- c("Gene1")

##load function to create density values
densMode <- function(x){
  td <- density(x)
  tdx <- td$x
  tdy <- td$y
  minx <- tdx[which(diff(sign(diff(tdy)))==2)]
  peakx <- tdx[which(diff(sign(diff(tdy)))==-2)]
  return(list(minx=minx, maxy=peakx))
}

## A density plot is made of the fds score, everything above a score of 0 is considered a doublet with another cell type
for (i in groups){
  ##create densityplots and set cut-offs
  vl <- densMode(fdscores@meta.data[,i])[1]
  vl2 <- densMode(fdscores@meta.data[,i])[2]
  plot(density(fdscores@meta.data[,i]), main=paste("densityplot, GenesetScore, The higher the cell's score, the higher the Geneset's average expression ",i, sep=""))
  if(density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] > 6 || density(fdscores@meta.data[,i])$y[which(density(fdscores@meta.data[,i])$x==vl$minx[1])] <0.01){
    abline(v=vl$minx[2])
    threshold=vl$minx[2]
  }else{
    abline(v=vl$minx[1], col="red")
    threshold=vl$minx[1]
  }
  #plot(hist(fdscores@meta.data[,i], plot=F, freq=FALSE))
  #abline(h=500)
  #abline(v=vl2$maxy, col="red")
  
  ##classify the cells based on thresholds of 0.1
  gbm.tumor_p53@meta.data[,paste("assignedto",i, sep="")] <- "nonclassified"
  gbm.tumor_p53@meta.data[which(fdscores@meta.data[,i]>0),paste("assignedto",i, sep="")] <- paste("assignedto",i, sep=" ")
}

## Doublets are removed
Idents(gbm.tumor_p53) <- gbm.tumor_p53@meta.data$assignedtoGene1
new.cluster.ids <- c( "Doublets", "Tumor cells")

names(new.cluster.ids) <- levels(gbm.tumor_p53)
gbm.tumor_p53 <- RenameIdents(gbm.tumor_p53, new.cluster.ids)
gbm.tumor_p53$celltype <- Idents(gbm.tumor_p53)

DimPlot(gbm.tumor_p53, reduction = "umap", group.by = "celltype")

## Clean up the data by removing doublets cells 
gbm.tumor_p53 <- gbm.tumor_p53[, gbm.tumor_p53$celltype %in% c("Tumor cells")]

DefaultAssay(gbm.tumor_p53) <- "integrated"
gbm.tumor_p53 <- RunUMAP(gbm.tumor_p53, reduction = "pca", dims = 1:30)
gbm.tumor_p53 <- FindNeighbors(gbm.tumor_p53, reduction = "pca", dims = 1:30)
gbm.tumor_p53 <- FindClusters(gbm.tumor_p53, resolution = 0.3, algorithm = 1)

DimPlot(gbm.tumor_p53, reduction = "umap", group.by = "seurat_clusters")
DimPlot(gbm.tumor_p53, reduction = "umap", group.by = "stage_model")
```


# Make gloioblastoma cellular subtpye (Neftel et al., 2019) mouse meta modules

```{r}
library(readxl)
library(nichenetr)

DefaultAssay(gbm.tumor_ink4a) <- "RNA"
DefaultAssay(gbm.tumor_p53) <- "RNA"

# Loading glioblastoma cellular subtype modules (Neftel et al., 2019)
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - code availibility/signatures"
setwd(work_dir)
Cell_Signatures <- read_excel("Glioblastoma_meta-module_genelist.xlsx")

MES1.features <- convert_human_to_mouse_symbols(Cell_Signatures$MES1)
MES2.features <- convert_human_to_mouse_symbols(Cell_Signatures$MES2)
AC.features <- convert_human_to_mouse_symbols(Cell_Signatures$AC)
OPC.features <- convert_human_to_mouse_symbols(Cell_Signatures$OPC)
NPC1.features <- convert_human_to_mouse_symbols(Cell_Signatures$NPC1)
NPC2.features <- convert_human_to_mouse_symbols(Cell_Signatures$NPC2)

## Function to assign cellular subtype
assignSubtype <- function (object, MES1.features,  MES2.features, AC.features, OPC.features, NPC1.features, NPC2.features ,set.ident = FALSE, ctrl_genes = 100,
                           ...)
{
  name <- "Cell.Subtype"
  features <- list(  MES1.Score = MES1.features, MES2.Score =  MES2.features, AC.Score = AC.features, OPC.Score = OPC.features, NPC1.Score = NPC1.features, NPC2.Score = NPC2.features)
  object.cc <- AddModuleScore(object = object,
                              features = features,
                              name = name,
                              ctrl = min(c(vapply(X = features, FUN = length, FUN.VALUE = numeric(length = 1))), 
                                         ctrl_genes), 
                              ...)
  cc.columns <- grep(pattern = name, x = colnames(x = object.cc[[]]),
                     value = TRUE)
  cc.scores <- object.cc[[cc.columns]]
  rm(object.cc)
  CheckGC()
  assignments <- apply(X = cc.scores, MARGIN = 1, FUN = function(scores,
                                                                 first = "MES1", second = "MES2", third = "AC", fourth = "OPC", fifth = "NPC1", sixth = "NPC2", null = "Undecided") {
    if (all(scores < -0)) {
      return(null)
    }
    else {
      return(c(first, second, third, fourth, fifth, sixth)[which(x = scores == max(scores))])
    }
  }
  )
  cc.scores <- merge(x = cc.scores, y = data.frame(assignments),
                     by = 0)
  colnames(x = cc.scores) <- c("rownames", "MES1.Score", "MES2.Score", "AC.Score","OPC.Score", "NPC1.Score", "NPC2.Score",
                               "Cell.Subtype")
  rownames(x = cc.scores) <- cc.scores$rownames
  cc.scores <- cc.scores[, c("MES1.Score", "MES2.Score", "AC.Score","OPC.Score", "NPC1.Score", "NPC2.Score","Cell.Subtype")]
  object[[colnames(x = cc.scores)]] <- cc.scores
  if (set.ident) {
    object[["old.ident"]] <- Idents(object = object)
    Idents(object = object) <- "Cell.Subtype"
  }
  return(object)
}

gbm.tumor_ink4a <- assignSubtype(gbm.tumor_ink4a, MES1.features = MES1.features, MES2.features = MES2.features, AC.features = AC.features, OPC.features = OPC.features, NPC1.features = NPC1.features, NPC2.features = NPC2.features)

gbm.tumor_p53 <- assignSubtype(gbm.tumor_p53, MES1.features = MES1.features, MES2.features = MES2.features, AC.features = AC.features, OPC.features = OPC.features, NPC1.features = NPC1.features, NPC2.features = NPC2.features)

# Plot cellular celltype distribution
DimPlot(gbm.tumor_ink4a, reduction = "umap", group.by = "Cell.Subtype", pt.size = 2)
DimPlot(gbm.tumor_p53, reduction = "umap", group.by = "Cell.Subtype", pt.size = 2)

# Merge data and create .csv file with subtype composition per samples
gbm.tumor_m <- merge(gbm.tumor_ink4a, gbm.tumor_p53)
x_abundance <- t(table(gbm.tumor_m$Cell.Subtype, gbm.tumor_m$samplename))
write.csv(x_abundance, file = "~/Desktop/x_abundance_Cell.Subtype-mouse-complete.csv")

# Prepare data
gbm.tumor_m <- gbm.tumor_m[, gbm.tumor_m$Cell.Subtype %in% c("AC", "MES1", "MES2", "NPC1", "NPC2", "OPC")]
gbm.tumor_m$MES.Score <- (gbm.tumor_m$MES1.Score + gbm.tumor_m$MES2.Score)/2
gbm.tumor_m$NPC.Score <- (gbm.tumor_m$NPC1.Score + gbm.tumor_m$NPC2.Score)/2

# D = max(SCopc,SCnpc) - max(SCac,SCmes)
gbm.tumor_m$max_OPC_NPC = ifelse(
  gbm.tumor_m$NPC.Score > gbm.tumor_m$OPC.Score, 
  gbm.tumor_m$NPC.Score, 
  gbm.tumor_m$OPC.Score)

gbm.tumor_m$max_AC_MES = ifelse(
  gbm.tumor_m$AC.Score > gbm.tumor_m$MES.Score, 
  gbm.tumor_m$AC.Score, 
  gbm.tumor_m$MES.Score)

gbm.tumor_m$D = gbm.tumor_m$max_OPC_NPC - gbm.tumor_m$max_AC_MES

gbm.tumor_m$x_axis <- ifelse(
  gbm.tumor_m$D > 0,
  log2(gbm.tumor_m$OPC.Score - gbm.tumor_m$NPC.Score+1),
  log2(gbm.tumor_m$AC.Score - gbm.tumor_m$MES.Score+1))


Idents(gbm.tumor_m) <- gbm.tumor_m$samplename

scatter <- subset(gbm.tumor_m, downsample = 500)
table(scatter@meta.data$samplename)
```


# Make representative cellular state plot per genetic background and stage

```{r}
## Give new ID - ink4a
DimPlot(gbm.tumor_ink4a, group.by = "Cell.Subtype", reduction = "umap")
Idents(gbm.tumor_ink4a) <- gbm.tumor_ink4a$Cell.Subtype
new.cluster.ids <- c("AC",	"MES",	"MES", "OPC", "NPC","NPC", "Undecided")
names(new.cluster.ids) <- levels(gbm.tumor_ink4a)
gbm.tumor_ink4a <- RenameIdents(gbm.tumor_ink4a, new.cluster.ids)
gbm.tumor_ink4a$Cell.Subtype.Grouped <- Idents(gbm.tumor_ink4a) 
gbm.tumor_ink4a <- gbm.tumor_ink4a[, gbm.tumor_ink4a$Cell.Subtype.Grouped %in% c("AC", "MES", "NPC", "OPC")]
DimPlot(gbm.tumor_ink4a, group.by = "Cell.Subtype.Grouped", reduction = "umap")

## Give new ID - p53
DimPlot(gbm.tumor_p53, group.by = "Cell.Subtype", reduction = "umap")
Idents(gbm.tumor_p53) <- gbm.tumor_p53$Cell.Subtype
new.cluster.ids <- c("AC",	"NPC",	"OPC", "MES", "MES","NPC", "Undecided")
names(new.cluster.ids) <- levels(gbm.tumor_p53)
gbm.tumor_p53 <- RenameIdents(gbm.tumor_p53, new.cluster.ids)
gbm.tumor_p53$Cell.Subtype.Grouped <- Idents(gbm.tumor_p53) 
gbm.tumor_p53 <- gbm.tumor_p53[, gbm.tumor_p53$Cell.Subtype.Grouped %in% c("AC", "MES", "NPC", "OPC")]
DimPlot(gbm.tumor_p53, group.by = "Cell.Subtype.Grouped", reduction = "umap")
```


# Figure 1C: Plot cellular subtype composition of each sample individually - PDG-Ink4a

```{r}
gbm.tumor <- gbm.tumor_ink4a
gbm.tumor$MES.Score <- (gbm.tumor$MES1.Score + gbm.tumor$MES2.Score)/2
gbm.tumor$NPC.Score <- (gbm.tumor$NPC1.Score + gbm.tumor$NPC2.Score)/2

# D = max(SCopc,SCnpc) - max(SCac,SCmes)
gbm.tumor$max_OPC_NPC = ifelse(
  gbm.tumor$NPC.Score > gbm.tumor$OPC.Score, 
  gbm.tumor$NPC.Score, 
  gbm.tumor$OPC.Score)

gbm.tumor$max_AC_MES = ifelse(
  gbm.tumor$AC.Score > gbm.tumor$MES.Score, 
  gbm.tumor$AC.Score, 
  gbm.tumor$MES.Score)

gbm.tumor$D = gbm.tumor$max_OPC_NPC - gbm.tumor$max_AC_MES

gbm.tumor$x_axis <- ifelse(
  gbm.tumor$D > 0,
  log2(gbm.tumor$OPC.Score - gbm.tumor$NPC.Score+1),
  log2(gbm.tumor$AC.Score - gbm.tumor$MES.Score+1))


Idents(gbm.tumor) <- gbm.tumor$samplename

scatter <- subset(gbm.tumor, downsample = 750)
table(scatter@meta.data$samplename)

## Colors based on non-grouped subtype
# PDG-ink4a model
ink4_prim1 <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_primary_CD45_b")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-Ink4a 1", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

ink4_prim2 <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_primary_CD45_c")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-Ink4a 2", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

ink4_prim3 <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_primary_CD45_d")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-Ink4a 3", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

ink4_rec1 <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_recurrent_CD45_b")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-Ink4a 1", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

ink4_rec2 <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_recurrent_CD45_c")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-Ink4a 2", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

ink4_rec3 <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_recurrent_CD45_d")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-Ink4a 3", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

in4a_prim <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_primary_CD45_b", "ink4a_primary_CD45_c", "ink4a_primary_CD45_d")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-Ink4a", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

ink4a_rec <- FeatureScatter(scatter[, scatter$samplename %in% c("ink4a_recurrent_CD45_b", "ink4a_recurrent_CD45_c", "ink4a_recurrent_CD45_d")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-Ink4a", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

# Figure 1C - Representative plot of glioblastoma cellular subtype heterogeneity within GEMM with PDG-Ink4a background 
in4a_prim ## Save PDF 12x4
```


# Plot cellular subtype composition of each sample individually - PDG-shp53 (not included in paper)

```{r}
gbm.tumor <- gbm.tumor_p53

gbm.tumor$MES.Score <- (gbm.tumor$MES1.Score + gbm.tumor$MES2.Score)/2
gbm.tumor$NPC.Score <- (gbm.tumor$NPC1.Score + gbm.tumor$NPC2.Score)/2

# D = max(SCopc,SCnpc) - max(SCac,SCmes)
gbm.tumor$max_OPC_NPC = ifelse(
  gbm.tumor$NPC.Score > gbm.tumor$OPC.Score, 
  gbm.tumor$NPC.Score, 
  gbm.tumor$OPC.Score)

gbm.tumor$max_AC_MES = ifelse(
  gbm.tumor$AC.Score > gbm.tumor$MES.Score, 
  gbm.tumor$AC.Score, 
  gbm.tumor$MES.Score)

gbm.tumor$D = gbm.tumor$max_OPC_NPC - gbm.tumor$max_AC_MES

gbm.tumor$x_axis <- ifelse(
  gbm.tumor$D > 0,
  log2(gbm.tumor$OPC.Score - gbm.tumor$NPC.Score+1),
  log2(gbm.tumor$AC.Score - gbm.tumor$MES.Score+1))


Idents(gbm.tumor) <- gbm.tumor$samplename

scatter <- subset(gbm.tumor, downsample = 750)
table(scatter@meta.data$samplename)

## Colors based on non-grouped subtype
# PDG-shp53 model
p53_prim1 <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_primary_CD45_a")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-p53 1", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_prim2 <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_primary_CD45_b")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-p53 2", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_prim3 <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_primary_CD45_c")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-p53 3", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_prim4 <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_primary_CD45_d")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary PDG-p53 4", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_rec1 <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_recurrent_CD45_a")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-Ink4a 1", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_rec2 <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_recurrent_CD45_b")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols =c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-Ink4a 2", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_rec3 <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_recurrent_CD45_c")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-Ink4a 3", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_prim <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_primary_CD45_a", "p53_primary_CD45_b", "p53_primary_CD45_c", "p53_primary_CD45_d")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols =c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Primary p53", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)

p53_rec <- FeatureScatter(scatter[, scatter$samplename %in% c("p53_recurrent_CD45_a", "p53_recurrent_CD45_c", "p53_recurrent_CD45_d")], feature1 = "x_axis", feature2 = "D", group.by = "Cell.Subtype", pt.size = 2, cols = c( "darkolivegreen3", "pink", "tomato3","darkgoldenrod1", "darkgoldenrod3", "royalblue3" )) + ggtitle("Recurrent PDG-p53", "Glioblastoma cellular subtype") + xlim(-0.5, 1) + ylim (-0.5, 0.7)
```


# Using IVY Gap scoring function to assess pseudolocation of tumor cells

```{r}
IVYGAPScoring <- function (object, ct.features,  pan.features, le.features, mvp.features ,set.ident = FALSE, ctrl_genes = 100,
              ...)
    {
        name <- "Cell.Cycle"
        features <- list(  PAN.Score = pan.features, LE.Score =  le.features, MVP.Score = mvp.features, CT.Score = ct.features)
        object.cc <- AddModuleScore(object = object,
                                    features = features,
                                    name = name,
                                    ctrl = min(c(vapply(X = features, FUN = length, FUN.VALUE = numeric(length = 1))), 
                                               ctrl_genes), 
                                    ...)
        cc.columns <- grep(pattern = name, x = colnames(x = object.cc[[]]),
                           value = TRUE)
        cc.scores <- object.cc[[cc.columns]]
        rm(object.cc)
        CheckGC()
        assignments <- apply(X = cc.scores, MARGIN = 1, FUN = function(scores,
                                                                       first = "PAN", second = "LE", third = "MVP", fourth = "CT", null = "Undecided") {
            if (all(scores < -0)) {
                return(null)
            }
            else {
                if (length(which(x = scores == max(scores))) > 1) {
                    return("Undecided")
                }
                else {
                    return(c(first, second, third, fourth)[which(x = scores == max(scores))])
                }
            }
        })
        cc.scores <- merge(x = cc.scores, y = data.frame(assignments),
                           by = 0)
        colnames(x = cc.scores) <- c("rownames", "PAN.Score", "LE.Score", "MVP.Score","CT.Score",
                                     "Location")
        rownames(x = cc.scores) <- cc.scores$rownames
        cc.scores <- cc.scores[, c("PAN.Score", "LE.Score", "MVP.Score" ,"CT.Score","Location")]
        object[[colnames(x = cc.scores)]] <- cc.scores
        if (set.ident) {
            object[["old.ident"]] <- Idents(object = object)
            Idents(object = object) <- "Location"
        }
        return(object)
    }
```


# Make tumor cell specific IVY Gap signatures

```{r}
library(nichenetr)
library(dplyr)

# Set working directory to folder "signatures" 
work_dir <- "~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - code availibility/signatures"
setwd(work_dir)

# Loading IVY_gap modules for pseudolocation function
IVY_gap <- read_excel("IVY_gap_signatures.xlsx")
IVY_gap$Gene <- IVY_gap$NAME %>% convert_human_to_mouse_symbols() 
IVY_gap <- IVY_gap[complete.cases(IVY_gap), ]

ct.all <-  subset(IVY_gap$Gene, subset = IVY_gap$Assigned == "CT")
pan.all <-  subset(IVY_gap$Gene, subset = IVY_gap$Assigned == "CTpan")
le.all <-  subset(IVY_gap$Gene, subset = IVY_gap$Assigned == "LE")
mvp.all <-  subset(IVY_gap$Gene, subset = IVY_gap$Assigned == "CTmvp")


ct.expression <- AverageExpression(gbm.combined_subset, assays = "RNA", features = ct.all, group.by = "celltype")
ct.expression <- ct.expression$RNA
ct.expression <- as.data.frame(ct.expression)

pan.expression <- AverageExpression(gbm.combined_subset, assays = "RNA", features = pan.all, group.by = "celltype")
pan.expression <- pan.expression$RNA
pan.expression <- as.data.frame(pan.expression)

le.expression <- AverageExpression(gbm.combined_subset, assays = "RNA", features = le.all, group.by = "celltype")
le.expression <- le.expression$RNA
le.expression <- as.data.frame(le.expression)

mvp.expression <- AverageExpression(gbm.combined_subset, assays = "RNA", features = mvp.all, group.by = "celltype")
mvp.expression <- mvp.expression$RNA
mvp.expression <- as.data.frame(mvp.expression)
```


# Using IVY Gap scoring function to assess pseudolocation of tumor cells

```{r}
library(ggplot2)
# Make tumor specific features
ct.features <- subset(rownames(ct.expression), subset = ct.expression$`Tumor cells` > 0.5)
pan.features <-  subset(rownames(pan.expression), subset = pan.expression$`Tumor cells` > 0.5)
le.features <-  subset(rownames(le.expression), subset = le.expression$`Tumor cells` > 0.5)
mvp.features <-  subset(rownames(mvp.expression), subset = mvp.expression$`Tumor cells` > 0.5)

DefaultAssay(gbm.tumor_ink4a) <- "RNA"
gbm.tumor_ink4a <- IVYGAPScoring(gbm.tumor_ink4a, ct.features = ct.features, pan.features = pan.features, le.features = le.features, mvp.features = mvp.features)
DimPlot(gbm.tumor_ink4a, reduction = "umap", group.by = "Location", label = F, pt.size = 2, cols = c("orange", "violetred1", "darkolivegreen3","lightblue", "grey") ) + ggtitle("Pseudolocation")
gbm.tumor_ink4a <- gbm.tumor_ink4a[, gbm.tumor_ink4a$Location %in% c( "MVP", "PAN", "CT", "LE")] 
gbm.tumor_ink4a$spatialsubtype <- paste(gbm.tumor_ink4a$Location, gbm.tumor_ink4a$Cell.Subtype)
Idents(gbm.tumor_ink4a) <- gbm.tumor_ink4a$Location
x_abundance_ink4a <- t(table(gbm.tumor_ink4a$spatialsubtype, gbm.tumor_ink4a$samplename))
write.csv(x_abundance_ink4a, file = "~/Desktop/x_gbm-sample-IVYGAP-ink4a.csv")

DefaultAssay(gbm.tumor_p53) <- "RNA"
gbm.tumor_p53 <- IVYGAPScoring(gbm.tumor_p53, ct.features = ct.features, pan.features = pan.features, le.features = le.features, mvp.features = mvp.features)
DimPlot(gbm.tumor_p53, reduction = "umap", group.by = "Location", label = F, pt.size = 2, cols = c("orange", "violetred1", "darkolivegreen3","lightblue", "grey") ) + ggtitle("Pseudolocation")
gbm.tumor_p53 <- gbm.tumor_p53[, gbm.tumor_p53$Location %in% c( "MVP", "PAN", "CT", "LE")] 
gbm.tumor_p53$spatialsubtype <- paste(gbm.tumor_p53$Location, gbm.tumor_p53$Cell.Subtype)
Idents(gbm.tumor_p53) <- gbm.tumor_p53$Location
x_abundance_p53 <- t(table(gbm.tumor_p53$spatialsubtype, gbm.tumor_p53$samplename))
write.csv(x_abundance_p53, file = "~/Desktop/x_gbm-sample-IVYGAP-p53.csv")
```


# Save polished tumor file

```{r}
# Set working directory to folder where integrated and filtered data will be saved: "scrna-seq"
save(gbm.tumor_p53,file="~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq/gbm-complete-tumor_p53.Rda")
save(gbm.tumor_ink4a,file="~/surfdrive/Code Availability/Kloosterman_et_al/Kloosterman et al. 2024 - data availibility/GEO Data - processed/scrna-seq//gbm-complete-tumor_ink4a.Rda")
```


# Figure 1D: Final quantification of pseudolocation across glioblastoma cellular subtype in murine dataset (data summarized as piechart Graphpad)

```{r}
x_abundance <- t(table(gbm.tumor$spatialsubtype, gbm.tumor$samplename)) ## Master table
write.csv(x_abundance, file = "~/Desktop/x_Pseudolocation-GBMsubtype_murine.csv")
```
